Repeat Yourself? When Linux kernel maintainers balance Code Duplication and Readability

### 2025-10-23 paper/techdebt2026/techdebt.tex
techdebt2026: techdebt: add takeaways w/ example quotes
From the patchset analysis, we detected 4 takeaways beyond the one
already in place. For now, dump them along with quotes in a temporary
section "Takeaways (temporary)", just to gauge the size of adding those
boxes. Beyond adding the takeaways and quotes, fill the column
"Takeaway" from the table with the contributions.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@

\begin{table}[ht]
\centering
\caption{Refactor methods used and metrics of the students' mitigations.}
\begin{tabular}{ |c |c |c |c |c | c| }
\hline
\textbf{GID} & \textbf{SS} & \textbf{SML} & \textbf{Patch Status} & \textbf{Takeaway} & \textbf{Diff} \\
\hline
0 & AMD & ???\% & Accepted (v3) & T1 & +114/-520 \\ \hline
1 & IIO & 100\% & Dropped (v2) & ? & +7/-14 \\ \hline
2 & IIO & 100\% & Dropped (v1) & ? & +23/-27 \\ \hline
3 & IIO & 100\% & Accepted (v1) & ? & +12/-30 \\ \hline
4 & IIO & 100\% & Accepted (v3) & ? & +2/-7 \\ \hline
5 & IIO & 90\% & Accepted (v1) & ? & +17/-70 \\ \hline
6 & IIO & 90\% & Dropped (v1) & ? & +95/-92 \\ \hline
7 & IIO & 90\% & Accepted (v3) & ? & +20/-36 \\ \hline
8 & IIO & 90\% & Accepted (v4) & ? & +20/-16 \\ \hline
9 & IIO & 90\% & Accepted (v6) & ? & +149/-243 \\ \hline
10 & IIO & 90\% & Dropped (v1) & ? & +32/-58 \\ \hline
11 & IIO & 100\% & Dropped (v1) & ? & +2/-12 \\ 
   & AMD & 90\% & Accepted (v2) & ? & +90/-489 \\ \hline

@@

>>>>

%% Takeaway Boxes %%
\usepackage[most]{tcolorbox}
\tcbset{colback=gray!10, colframe=gray!50!black, boxrule=0.5pt,
        left=1mm, right=1mm, top=1mm, bottom=1mm, sharp corners}

\newenvironment{takeaway}[1]{%
  \begin{tcolorbox}
  \textbf{#1:} \itshape}{\end{tcolorbox}}
%% Takeaway Boxes %%

@@

\begin{table}[ht]
\centering
\caption{Refactor methods used and metrics of the students' mitigations.}
\begin{tabular}{ |c |c |c |c |c | c| }
\hline
\textbf{GID} & \textbf{SS} & \textbf{SML} & \textbf{Patch Status} & \textbf{Takeaway} & \textbf{Diff} \\
\hline
0 & AMD & ???\% & Accepted (v3) & T1 & +114/-520 \\ \hline
1 & IIO & 100\% & Dropped (v2) & T2, T3 & +7/-14 \\ \hline
2 & IIO & 100\% & Dropped (v1) & T2 & +23/-27 \\ \hline
3 & IIO & 100\% & Accepted (v1) & None & +12/-30 \\ \hline
4 & IIO & 100\% & Accepted (v3) & T2 & +2/-7 \\ \hline
5 & IIO & 90\% & Accepted (v1) & None & +17/-70 \\ \hline
6 & IIO & 90\% & Dropped (v1) & T4 & +95/-92 \\ \hline
7 & IIO & 90\% & Accepted (v3) & None & +20/-36 \\ \hline
8 & IIO & 90\% & Accepted (v4) & T5 & +20/-16 \\ \hline
9 & IIO & 90\% & Accepted (v6) & T5 & +149/-243 \\ \hline
10 & IIO & 90\% & Dropped (v1) & T2, T5 & +32/-58 \\ \hline
11 & IIO & 100\% & Dropped (v1) & None & +2/-12 \\
   & AMD & 90\% & Accepted (v2) & None & +90/-489 \\ \hline

@@

%TODO REMOVE ME!!! Essa seção é só para fazer o dump e ver como fica o tamanho.
\section{Takeaways (temporary)}

%TODO achar uma forma mais visual de colocar as aspas
\noindent
\begin{footnotesize}
  \textit{
    Quote: ``I think the old code is more readable than hiding the values in a macro
    even if it is duplicating a few lines of code.''
  }
\end{footnotesize}

\begin{takeaway}{Takeaway T2 (Readability)}
  Maintainers notably prioritized readability over removing duplications in
  many different contexts. In their view, going against the \textbf{DRY
  principle} was worth it if it resulted in code that is easier to read and
  understand.
\end{takeaway}

\noindent
\begin{footnotesize}
\textit{
    Quote: ``[...] such patches might not worth it since the proposed
    improvement is very small (and questionable) while the upstreaming process
    still requires some effort.''
  }
\end{footnotesize}

\begin{takeaway}{Takeaway T3 (Integration Overhead)}
  A maintainer suggested that the effort put into the contribution after it was
  accepted (i.e., the process of propagating the change to upper maintainers)
  was considered. In other words, maintainers may \textbf{evaluate the
  trade-off between the impact of contributions (not only deduplications) and
  the struggle to integrate them into the complex Linux project}.
\end{takeaway}

\noindent
\begin{footnotesize}
\textit{
    Quote: ``Wrapping this up doesn't provide any real advantage, requiring as
    it does the reviewer to look at this function AND where the value is set
    rather than seeing them in one place.''
  }
\end{footnotesize}

%TODO Na citação abaixo, trazida pelo Arthur, pensar se não vale colocar as
%referências base do paper do Lanza: ``Developers need to mentally reconstruct
%the role and behavior of a class from scattered fragments of code [8],
%presented as one page of text at a time. This process is cognitively demanding
%[9], slow, and error-prone, especially in large systems with many
%interdependent components [10].''
%
% Referências do paper do Lanza para esse trecho:
%[8] M.-A. Storey, “Theories, methods and tools in program comprehension:
%Past, present and future,” in Proceedings of IWPC 2005. IEEE, 2005,
%pp. 181–191.
%[9] J. Sillito, G. C. Murphy, and K. De Volder, “Questions programmers
%ask during software evolution tasks,” in Proceedings FSE 2006. ACM,
%2006, pp. 23––34.
%[10] G. C. Murphy, D. Notkin, and K. Sullivan, “Software reflexion models:
%Bridging the gap between source and high-level models,” in Proceedings
%of FSE 1995. ACM, 1995, pp. 18–28.
\begin{takeaway}{Takeaway T4 (Cognitive Load)}
  Duplications that made it possible to read the code without needing to jump
  to the function definition (to see its behavior) or the function call (to see
  its concrete parameters) were justifiable to reduce cognitive load when
  parsing through the code~\cite{skylines-cognitive-load}.
\end{takeaway}

\noindent
\begin{footnotesize}
\textit{
    Quote: ``Note that is not an appropriate change for the large reads though
    as \texttt{spi\_write\_then\_read()} bounces all buffers and so would add a
    copy to those high(ish) performance paths.''
  }
\end{footnotesize}

\begin{takeaway}{Takeaway T5 (Performance)}
  Given the low-level development context of Linux and the IIO subsystem, which
  is usually related to devices with limited resources, we detected hints that
  duplication may be forgivable if it improves performance.
\end{takeaway}
